---
title: "R_project"
output: html_document
date: "2025-01-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Super Bowl Ad Engagement Analysis

## Introduction

Super Bowl advertisements are a crucial aspect of the annual sports event, attracting millions of viewers and significant brand investment. This project analyzes Super Bowl ad engagement metrics to determine what factors contribute to higher audience interactions, particularly through YouTube likes and comments. The study also explores brand performance, humor's impact on engagement, and the relationship between TV viewership and online interactions.

## Dataset Overview

The dataset consists of Super Bowl advertisements from 2000 onwards, enriched with additional metrics such as:

  - Engagement Metrics: Likes, dislikes, comments, and views on YouTube
  - Attributes: Whether the ad features humor, celebrities, animals, patriotic themes, danger, or explicit product placement  
  - TV Metrics: TV views in millions, NFL revenue from advertisements
  - Ad Descriptions: Brand, year, ad url

The data was sourced from Tidy Tuesday data on GitHub and enriched using statista.

⚠ Dataset Limitation:
This dataset does not include all Super Bowl ads. Instead, it consists of 233 ads from the 10 brands that aired the most frequently during Super Bowls in the past two decades. The data was sourced from SuperBowlAds.com, meaning some brands and ads are missing.

```{r imporitng libraries and data}
library(data.table)
library(skimr)
library(readr)
library(tidyr)
library(ggplot2)
library(scales)
library(RColorBrewer)

#superbowl data from tody tuesday 
sb <- fread('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2021/2021-03-02/youtube.csv')
sb <- as.data.table(sb)
skim(sb)

#enriching the data by scraping data from Statista
superbowl_scraped <- fread("https://raw.githubusercontent.com/ghadena/C3-R/refs/heads/main/superbowl_scraped.csv")
superbowl_scraped <- as.data.table(superbowl_scraped)
superbowl_scraped <- superbowl_scraped[year >= 2000, ] #filtering for years after 2000 
skim(superbowl_scraped)
```

## Research Questions

1. What attributes in Super Bowl ads contribute to higher engagement?

Examining the impact of humor, celebrity appearances, patriotism, and other attributes on likes and comments.

2. Why do funny ads get fewer likes and comments?

Investigating whether humor interacts with other factors to influence engagement.

3. Which brands consistently get the highest engagement, and why?

Analyzing top advertisers based on engagement and identifying key patterns.

4. How does TV viewership impact online engagement?

Assessing whether years with higher TV views correspond to higher YouTube likes.

## Data Processing and Cleaning

Filtered dataset to include only years 2000 onwards.

Merged additional data sources containing TV views and NFL ad revenue.

Normalized brand engagement metrics to compare advertisers fairly.

Handled missing values by omitting NA values from calculations.

Converted categorical attributes to a long format for better visualization.

```{r preparing data}

# Merging 
dt <- merge(sb, superbowl_scraped, by = "year")

# Counting commercials by brand 
sb[ , .N, by = .(brand)][order(-N)] 

# Convert logical variables to long format
logical_vars <- sb %>%
  pivot_longer(cols = where(is.logical), names_to = "Variable", values_to = "Value")

```
## Attribute Analysis: What Makes Ads More Engaging?

We examined seven key attributes across Super Bowl commercials:
Humor, Celebrities, Animals, Patriotism, Danger, Sex Appeal and Product Placement

To answer the first research question, I aggregated the dataset, calculating the average likes and average comments for ads with and without each attribute. Then counted how many times each attribute appeared in the dataset to understand their prevalence. and finally created visualizations to compare engagement across different categories.

Findings:
	**- Patriotic ads**: While most Super Bowl ads do not feature patriotic imagery, those that do receive the highest average likes and comments, possibly due to their emotional appeal.
	**- Product visibility:** Many ads show the product quickly, but this does not significantly affect likes or comments.
	**- Use of sex:** The majority of ads avoid using sexual content—likely because the Super Bowl is a family-friendly event.
	**-	Celebrity appearances & danger elements:** Ads featuring celebrities and a sense of danger tend to generate more likes and comments than those without these elements.
	**-	Humorous ads paradox:** Surprisingly, funny ads received fewer likes and comments compared to non-funny ads. Initially, we suspected this was due to a lower number of funny ads, but this was not the case—there were almost twice as many funny ads as non-funny ones. This unexpected trend led us to investigate further in the next section.


```{r Attribute analysis}

# Define the list of categorical columns
categories <- c("funny", "patriotic", "celebrity", "danger", "animals", "use_sex", "show_product_quickly")

# Initialize an empty data.table to store aggregated results
cat_agg <- data.table(category = character(), value = logical(), avg_likes = numeric(), avg_comments = numeric())

# Loop through each category and compute avg_likes and avg_comments
for (category in categories) {
  temp_dt <- dt[, .(avg_likes = mean(like_count, na.rm = TRUE), 
                    avg_comments = mean(comment_count, na.rm = TRUE)), 
                by = category]
  setnames(temp_dt, category, "value") # Rename the grouping column to a generic name "value"
  temp_dt[, category := category]  # Add a new column to indicate the category name
  cat_agg <- rbind(cat_agg, temp_dt, fill = TRUE)   # Bind the results to the main results_dt
}
setcolorder(cat_agg, c("category", "value", "avg_likes", "avg_comments")) # Reorder columns for clarity
print(cat_agg)

```
Visualizations

To illustrate these insights, I included three grouped bar charts:
	1.	Count of occurrences of each attribute in the dataset.
	2.	Average likes received by ads based on whether they had the attribute or not.
	3.	Average comments received by ads based on whether they had the attribute or not.

These visualizations help us identify patterns in Super Bowl ad engagement and provide insights into which elements contribute to higher audience interaction.
```{r visuals}

#chart 1 counting the occurrence of categories 
ggplot(logical_vars, aes(x = Variable, fill = Value)) +
  geom_bar(position = "dodge") +
  ggtitle("Count of Ad Attributes") +
  xlab("Categorical Attributes") +
  ylab("Count") +
  theme_minimal() +
  scale_fill_manual(values = c("TRUE" = "blue", "FALSE" = "red"))

#chart 2 average likes by category 
ggplot(cat_agg, aes(x = category, y = avg_likes, fill = value)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Average Likes by Category", x = "Category", y = "Average Likes", fill = "Value (TRUE/FALSE)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#chart 3 average commnets by category
ggplot(cat_agg, aes(x = category, y = avg_comments, fill = value)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Average Comments by Category", x = "Category", y = "Average Comments", fill = "Value (TRUE/FALSE)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

## Does Humor Really Boost Engagement? Investigating Funny Super Bowl Ads

My initial analysis showed that funny ads received fewer likes and comments than non-funny ones, despite making up the majority of the dataset. To understand why, I explored several factors.

**Do Funny Ads Actually Get Less Engagement?**

Since humor was expected to drive engagement, I first checked whether funny ads simply had fewer observations, but they actually outnumbered non-funny ads nearly 2:1. This ruled out sample size as a factor. Next, I examined whether funny ads had more variation in likes, but the standard deviation was lower than for non-funny ads. This means funny ads were more consistent in performance. I also analyzed the trend of average likes over time for funny vs. non-funny ads, but the results were inconclusive, showing a largely random pattern—possibly due to a few viral ads and small sample size heavily distorting the data.

**Do Other Factors Influence Funny Ads’ Performance?**

Certain ad attributes, like celebrity appearances, typically boost engagement. However, I found that celebrity ads that weren’t funny performed best (~10,000 likes on average), while funny ads without celebrities got ~4,700 likes. Funny ads with celebrities performed even worse. This suggests that while celebrities usually increase engagement, humor weakens their impact, possibly by shifting focus away from them.

**Are Funny Ads More Entertaining But Less Engaging?**

Although funny ads received fewer likes per view, they were watched more often than non-funny ads. This suggests they are entertaining enough to keep viewers watching, but not compelling enough to drive interaction. A check on dislikes revealed that funny ads also received fewer dislikes, reinforcing the idea that they are well-received but not particularly engaging.

**Are Funny Ads Misclassified?**

The dataset defines a funny ad as jokey, goofy, weird, or silly, or one that attempts to be humorous. However, after reviewing top-performing ads, I found that some—like Super Bowl Babies Choir and Budweiser’s Donkey ad—weren’t labeled as funny, even though they fit the definition. This suggests potential misclassification, which may affect results. 

Humor is highly subjective—what one person finds funny, another may not. Unlike other attributes like celebrity appearances, animals, or product placement, which are more objective, humor varies by perception. This subjectivity could explain why the data contradicts expectations, showing that funny ads receive fewer likes and comments despite their entertainment value.

```{r funny ads}
#lets check the distribution of funny videos (count and standard deviation)
dt[, .N, by = funny] 
dt[, .(avg_likes = mean(like_count, na.rm = TRUE), 
       sd_likes = sd(like_count, na.rm = TRUE)), 
   by = funny] 

#examining the trend overtime 
dt[, .(avg_likes = mean(like_count, na.rm = TRUE)), by = .(year, funny)][order(year, -avg_likes)]
ggplot(dt[, .(avg_likes = mean(like_count, na.rm = TRUE)), by = .(year, funny)], 
       aes(x = year, y = avg_likes, color = funny, group = funny)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(title = "Trend of Average Likes Over Time (Funny vs. Not Funny Ads)", 
       x = "Year", y = "Average Likes", color = "Funny") +
  theme_minimal()

#examining the influence of other factors
dt[, .(avg_likes = mean(like_count, na.rm = TRUE)), by = .(funny, celebrity, animals)][order(-avg_likes)] 

#funny videos get more views but less likes 
dt[, .(avg_views = mean(view_count, na.rm = TRUE), avg_likes = mean(like_count, na.rm = TRUE)), by = funny] 
dt[, likes_per_view := like_count / view_count]
ggplot(dt[, .(avg_likes_per_view = mean(likes_per_view, na.rm = TRUE)), by = funny], 
       aes(x = as.factor(funny), y = avg_likes_per_view, fill = as.factor(funny))) +
  geom_bar(stat = "identity") +
  labs(title = "Likes per View: Funny vs. Not Funny Ads", 
       x = "Funny", y = "Average Likes per View") +
  theme_minimal()   

#lets check the another engagement metric, number of dislikes.
dt[, .(avg_dislikes = mean(dislike_count, na.rm = TRUE)), by = funny]

#lets check the top 20 videos by engagement and examine potential classification issues. (superbowl babay choir and the budwiser donley videos are misclassified imo)
dt[order(-like_count)][1:20, .(year, funny, brand, title, like_count, view_count)]

```
**Visualizing Likes Distribution for Funny vs. Non-Funny Ads**

To compare engagement, I created three boxplots:
	•	The first included all ads, showing non-funny ads had higher likes overall.
	•	The second excluded the top 5% of extreme values, revealing clearer patterns.
	•	The third used a log scale, further confirming that non-funny ads consistently received more likes than funny ones.
```{r funny ads visuals}

# including all observations
ggplot(dt, aes(x = as.factor(funny), y = like_count, fill = as.factor(funny))) +
  geom_boxplot() +
  labs(title = "Likes Distribution: Funny vs. Non-Funny Ads", 
       x = "Funny", y = "Like Count", fill = "Funny") +
  theme_minimal()

# excluding outliars - top 5% 
dt_filtered <- dt[like_count <= quantile(like_count, 0.95, na.rm = TRUE)]

ggplot(dt_filtered, aes(x = as.factor(funny), y = like_count, fill = as.factor(funny))) +
  geom_boxplot() +
  labs(title = "Likes Distribution (Excluding Top 5% Outliers): Funny vs. Non-Funny Ads", 
       x = "Funny", y = "Like Count", fill = "Funny") +
  theme_minimal()

# on log scale 
ggplot(dt, aes(x = as.factor(funny), y = like_count, fill = as.factor(funny))) +
  geom_boxplot() +
  scale_y_log10() +
  labs(title = "Likes Distribution (Log Scale): Funny vs. Non-Funny Ads", 
       x = "Funny", y = "Log10(Likes)", fill = "Funny") +
  theme_minimal()

```

**Ad Success Is More Than Just Likes**

While my analysis focuses on likes and comments as engagement metrics, they do not fully define ad success. A truly successful ad could drive sales, brand awareness, or cultural impact—factors that are difficult to quantify but often more important than online engagement. Some funny ads may generate buzz or brand recognition without necessarily earning more likes.

While humor is often seen as a winning strategy, simply being funny does not guarantee engagement. Other elements—emotional appeal, storytelling, and branding strategies—may play a bigger role in making an ad truly successful.

## Brand Performance: Which Brands Dominate Engagement?

To investigate which brands consistently received higher likes and engagement, I created two datasets:
	1.	Brand Performance Table – Calculated average likes, comments, dislikes, and ad counts for each brand.
	2.	Attribute Composition Table – Measured how often each brand used specific features (e.g., funny, celebrity, patriotic) in their ads, expressed as percentages. This table also included average likes, comments, dislikes, and views per brand.
	
	
```{r brand enagement }
# Calculate average likes, comments, dislikes, and ad counts for each brand
brand_engagement <- dt[, .(
  avg_likes = mean(like_count, na.rm = TRUE),
  avg_comments = mean(comment_count, na.rm = TRUE),
  avg_dislikes = mean(dislike_count, na.rm = TRUE),
  total_ads = .N
), by = brand][order(-avg_likes)]

print(brand_engagement)

# Calculate attribute contribution and average engagement (likes, comments, dislikes)
attribute_contribution <- dt[, .(
  funny_pct = mean(funny, na.rm = TRUE) * 100,
  celebrity_pct = mean(celebrity, na.rm = TRUE) * 100,
  animals_pct = mean(animals, na.rm = TRUE) * 100,
  product_pct = mean(show_product_quickly, na.rm = TRUE) * 100,
  patriotic_pct = mean(patriotic, na.rm = TRUE) * 100,
  danger_pct = mean(danger, na.rm = TRUE) * 100,
  sex_pct = mean(use_sex, na.rm = TRUE) * 100,
  avg_likes = mean(like_count, na.rm = TRUE),
  avg_comments = mean(comment_count, na.rm = TRUE),
  avg_dislikes = mean(dislike_count, na.rm = TRUE), 
  avg_views = mean(view_count, na.rm = TRUE)  # Removed * 100
), by = brand][order(-avg_likes)]

print(attribute_contribution)
```

**Which Brands Receive the Most Likes?**

A simple bar chart of average likes per brand revealed that:
	•	NFL led with ~25,000 average likes per ad.
	•	Doritos followed with ~15,000.
	•	Coca-Cola, Budweiser, and Bud Light came next.
	•	The remaining brands—Pepsi, Toyota, E-Trade, Hyundai, and Kia—all averaged under 1,000 likes per ad.
	
```{r avg likes}
#average likes per brand 
ggplot(attribute_contribution, aes(x = reorder(brand, -avg_likes), y = avg_likes, fill = brand)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Likes by Brand", x = "Brand", y = "Average Likes") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

**What Makes Certain Brands More Engaging?**

To explore why some brands performed better, I normalized attribute percentages per brand to sum to 100% and converted the data into a long format for visualization.
```{r visuals2}
# Normalize attribute percentages to sum to 100% per brand
attribute_contribution[, total_pct := funny_pct + celebrity_pct + animals_pct + product_pct + patriotic_pct + danger_pct + sex_pct]
attribute_contribution[, ':=' (
  funny_pct = funny_pct / total_pct * 100,
  celebrity_pct = celebrity_pct / total_pct * 100,
  animals_pct = animals_pct / total_pct * 100,
  product_pct = product_pct / total_pct * 100,
  patriotic_pct = patriotic_pct / total_pct * 100,
  danger_pct = danger_pct / total_pct * 100,
  sex_pct = sex_pct / total_pct * 100
)]

# Convert data to long format for the stacked bar
attribute_long <- melt(attribute_contribution, id.vars = "brand",
                       measure.vars = c("funny_pct", "celebrity_pct", "animals_pct", "product_pct", "patriotic_pct", "danger_pct", "sex_pct"),
                       variable.name = "attribute", value.name = "percentage")

```

A stacked bar chart showed that NFL, the top performer, used celebrities the most, which may have contributed to its high engagement. NFL ads also incorporated patriotic themes more than any other brand, suggesting that emotional appeal might play a role in driving likes.

```{r stacked bar chart}
# chart 6: Create a 100% stacked bar chart
ggplot(attribute_long, aes(x = reorder(brand, -percentage), y = percentage, fill = attribute)) +
  geom_bar(stat = "identity", position = "fill") +  # "fill" ensures each bar sums to 100%
  scale_y_continuous(labels = scales::percent_format(scale = 100)) +  # Format y-axis as percentages
  labs(title = "Normalized Attribute Usage by Brand (100% Stacked Bar Chart)", 
       x = "Brand", y = "Percentage of Ads", fill = "Attribute") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

**Adding an Extra Layer: Likes as a Percentage of Total Engagement**

To enhance the analysis, I calculated each brand’s share of total likes and overlaid this on the stacked bar chart. This provided a clearer picture of how brand attributes correlate with engagement levels.

Overall, brands that leveraged emotional appeal—celebrities, patriotism, and nostalgia—tended to achieve higher engagement, while those that relied more on product-focused or action-driven ads performed less consistently.

```{r stacked bar chart w/ percentage of likes}
# Calculate % of total likes for each brand
total_likes <- sum(attribute_contribution$avg_likes, na.rm = TRUE)
attribute_contribution[, likes_pct := (avg_likes / total_likes) * 100]

ggplot() +
  # Stacked bar chart (attributes)
  geom_bar(data = attribute_long, 
           aes(x = brand, y = percentage / 100, fill = attribute), 
           stat = "identity", position = "fill") +
  
  # Line and points for % of total likes (on the same y-axis)
  geom_line(data = attribute_contribution, 
            aes(x = brand, y = likes_pct / 100, group = 1, color = "Likes % of Total"), 
            size = 1.5) + 
  geom_point(data = attribute_contribution, 
             aes(x = brand, y = likes_pct / 100, color = "Likes % of Total"), 
             size = 2) +  
  scale_color_manual(values = c("Likes % of Total" = "black"), name = "Metric") +  # Add to legend
  
  # Labels for % of total likes
  geom_text(data = attribute_contribution, 
            aes(x = brand, y = (likes_pct / 100) + 0.04, label = paste0(round(likes_pct, 1), "%")), 
            vjust = -0, color = "black", size = 2.5) +
  
  # Formatting
  scale_y_continuous(labels = percent_format(), 
                     breaks = seq(0, 1, by = 0.25),  # Set ticks at 0, 25, 50, 75, 100
                     expand = expansion(mult = c(0, 0.1))) +  # Add space for labels
  
  labs(title = "Super Bowl Ad Strategies: Attribute Breakdown and Engagement Comparison by Brand", 
       subtitle = "Analyzing the attribute composition of Super Bowl ads by top brands \nand how their content translates into audience engagement (measured by percentage of total likes).",
       x = "Brand", y = "Percentage", fill = "Attribute") +
  
  # Theme adjustments
  theme_minimal(base_size = 12) +  # Keep minimal theme
  theme(panel.grid.major = element_blank(),   # Remove major grid lines
        panel.grid.minor = element_blank(),   # Remove minor grid lines
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        plot.title = element_text(hjust = 0.5, size = 15, margin = margin( b = 10)),  # Add top margin
        plot.subtitle = element_text(hjust = 0, size = 10, margin = margin(b = 5)),  # Add bottom margin
        plot.margin = margin(t = 20, r = 0, b = 10, l = 10)
        )  # Better x-axis rotation
```

## Does Higher TV Viewership Lead to More Online Engagement?

One question I explored was whether years with higher Super Bowl TV viewership also led to more engagement on YouTube. If more people saw the ads on TV, did that translate into more online reactions? This could help determine whether traditional TV exposure still influences digital engagement or if some brands are simply better at making viral ads.

**Data Challenges & Adjustments**

Initially, I plotted a scatter plot of TV views vs. YouTube likes per ad, but the visualization was misleading because:
	1.	TV viewership data was aggregated per year, while likes were per individual ad, making direct comparisons inaccurate.
	2.	A single extreme outlier—the 2012 Doritos ad (275,000 likes)—heavily skewed the trend.
	
```{r TV graph 1}
#correcting column name
setnames(dt, "tv_views(millions)", "tv_views_million")

#TV viewership and youtube likes 
ggplot(dt, aes(x = tv_views_million, y = like_count)) +
  geom_point(alpha = 0.6) +
  geom_point(data = dt[which.max(like_count)], aes(x = tv_views_million , y = like_count), 
             color = "red", size = 3) +
  scale_x_continuous(labels = scales::comma) +  # No scientific notation
  scale_y_continuous(labels = scales::comma) +  # No scientific notation
  labs(title = "YouTube Views vs. Likes",
       x = "TV Views (millions)", y = "Likes",
       subtitle = "Red point = Outlier (Doritos 2012 ad) with 275k Youtube Likes") +
  theme_minimal()

#identifying the outliar
dt[which.max(view_count)]
dt[which.max(like_count)]

```

To correct this, I aggregated likes per year and created a new yearly engagement dataset, allowing for a clearer comparison between average YouTube likes and total TV viewership per year.

```{r annotaged scatterplot}
# Aggregate data: Average likes per year
yearly_engagement <- dt[, .(avg_likes = mean(like_count, na.rm = TRUE)), by = .(year, tv_views_million)]

# Load necessary library for ColorBrewer palettes
display.brewer.all()  # Shows all available palettes

# Define a color palette for years
year_colors <- brewer.pal(n = 12, name = "RdPu")  # Using a ColorBrewer palette

# Identify the outliers
max_outlier <- yearly_engagement[which.max(avg_likes)]  # 2012 (doritos ad - 275k likes)
second_outlier <- yearly_engagement[order(-avg_likes)][2]  # 2020 
second_outlier

dt[order(-like_count)][2]  # (nfl LIV comercial - 175k likes)

# Scatter plot with color and annotations
ggplot(yearly_engagement, aes(x = tv_views_million, y = avg_likes, color = year)) +
  geom_point(size = 4) +  # Bigger points
  geom_smooth(method = "lm", se = FALSE, color = "black", linetype = "dashed") +  # Trend line
  scale_x_continuous(labels = scales::comma) +  
  scale_y_continuous(labels = scales::comma) +  
  scale_color_gradientn(colors = brewer.pal(n = 9, name = "RdPu")) +  # Apply color scheme
  labs(
    title = "TV Views vs. Average YouTube Likes per Year",
    subtitle = "Analyzing whether years with more Super Bowl viewers generate more online engagement",
    x = "TV Views (millions)",
    y = "Average YouTube Likes",
    color = "Year",
    caption = "Each point represents a year. Trend line shows overall correlation."
  ) +
  theme_minimal() +
  
  # Annotation for Max Outlier (Doritos 2012 Ad)
  annotate("text", x = max_outlier$tv_views_million, y = max_outlier$avg_likes + 5000,
           label = "Doritos 2012 Ad\n~275k Likes", color = "black", size = 4, hjust = 0.5) +
  geom_segment(aes(x = max_outlier$tv_views_million, y = max_outlier$avg_likes - 20000,
                   xend = max_outlier$tv_views_million, yend = max_outlier$avg_likes - 5000),
               arrow = arrow(length = unit(0.2, "cm")), color = "black") +
  
  # Annotation for Second Outlier
  annotate("text", x = second_outlier$tv_views_million, y = second_outlier$avg_likes + 5000,
           label = paste0("NFL 2020 Ad\n~178k Likes"), 
           color = "black", size = 4, hjust = 0.5) +
  geom_segment(aes(x = second_outlier$tv_views_million, y = second_outlier$avg_likes - 20000,
                   xend = second_outlier$tv_views_million, yend = second_outlier$avg_likes - 5000),
               arrow = arrow(length = unit(0.2, "cm")), color = "black")

```

**Findings: The Impact of TV Exposure on YouTube Likes**

The revised scatter plot provided more context by specifically highlighting two major outliers:
	•	2012 Doritos Ad (275K likes)
	•	2020 NFL LIV Commercial (175K likes)

Additionally, the visualization used a ColorBrewer palette to differentiate the years, making it easier to spot trends over time. This helped reveal a positive correlation between TV views and online engagement, suggesting that in years with higher TV viewership, ads also received more YouTube likes. However, this trend was heavily influenced by the two outliers, indicating that virality, rather than TV exposure alone, may be a key driver of engagement.

**Possible Explanations**
	1.	Changing Viewing Habits & Online Presence
	•	Over the years, people have become more active online, leading to higher engagement rates regardless of TV viewership.
	•	The increase in social media usage and YouTube popularity likely contributes to the rising trend in likes over time.
	2.	How People Watch the Super Bowl Has Changed
	•	In the early 2000s, many people watched the game in sports bars or large group settings, where engagement with ads may have been lower.
	•	Today, many viewers watch individually on personal devices and engage more actively by liking and sharing ads online.
	•	Some may even skip TV ads altogether and watch the best commercials later on YouTube.
	3.	Platform Shifts: The TikTok & Instagram Factor
	•	Engagement patterns may be changing, with younger audiences shifting from YouTube to platforms like TikTok and Instagram.
	•	Companies may now focus more on short-form ad placements, which could explain why YouTube likes alone may not capture the full picture of ad success.

While TV exposure may still play a role in ad success, online engagement is shaped by many other factors, from virality and content strategy to shifts in audience behavior.

## Conclusion and Key Takeaways:

- Humor alone does not guarantee high engagement – combining humor with celebrities may reduce effectiveness.
- Celebrity endorsements and patriotic themes tend to increase engagement, particularly in NFL and Pepsi ads.
- NFL commercials dominate online engagement, likely due to audience interest and familiarity with the brand.
- Higher TV views do not always translate into higher YouTube engagement, indicating that viral success is independent of broadcast reach.
- Certain years, such as 2012 (Doritos) and 2020 (NFL), skew the trend due to highly viral ads.

**Future Considerations:**

- Expanding engagement analysis beyond YouTube (e.g., Twitter, TikTok, and sales data).
- Investigating whether certain ad types lead to increased product sales.
- Analyzing sentiment analysis from YouTube comments to understand audience reactions.


Prepared by Ghadena Hgaig 

January 2025

